---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

# introduction

this scripts contains code used to plot the dna segments from the assemblies

# installing packages
```{r, eval=FALSE}

# install.packages("tidyverse") 
# install.packages("genoPlotR")
# install.packages("janitor")
# install.packages("stringr")
# install.packages("config")
#remotes::install_github("clauswilke/colorblindr")
```

# the chromosome
## loading files

```{r}
# load the packages
library(genoPlotR)
library(tidyverse)
library(janitor)
library(patchwork)
library(ragg)
library(forcats)
library(RColorBrewer)
library(viridis)
library(colorblindr) #the packge has a function {cvd_grid(plot)} 
#which can simulate how a color blind person would see the plots

setwd("/Users/zuza/repos/H58")

path_to_genbank_files <- "outputs/gbk"

sample_type <- "chromosome"

file_names <- 
c("1017142", "BKQU3X", "A58390", "BKQT8S", "AL513382.1") # removed "A53789"

extension <- ".gbk"

# load blast comparisons
comparison_file_1 <-
  "outputs/blast/ct18_vs_1017142.txt"
comparison_file_2 <-
  "outputs/blast/1017142_vs_A58390.txt"
comparison_file_3 <-
  "outputs/blast/A58390_vs_BKQT8S.txt"
comparison_file_4 <-
  "outputs/blast/BKQT8S_vs_BKQU3X.txt"


# load the phage regions in the CT18 ref genome
phage_data_location <-"outputs/phaster"

# create custom color pellet
colors <- 
    unique(c(brewer.pal(n = 11, name = "Paired"), brewer.pal(n = 8, name = "Dark2"), brewer.pal(n = 8, name = "Set2")))

colors <- viridis(n = 20)
# output folder

outdir <- "outputs/plots"
```


# bacteriophage region positions in the reference
Using phaster we located bacteriophages on the chromosome. 
these regions are plotted over the next few lines. a total of 8 phage regions were identified and 5 of them are intact

### Phage 1
a 45.7kb intact phage spanning the region 1007350-1053060 of the ct18 reference genome.


```{r }
sample_name <- str_glue(file_names[5], extension)

ct18 <- file.path(path_to_genbank_files, sample_type, sample_name)
  
dna_seg <-
  read_dna_seg_from_genbank(ct18)


  # which gets the row number of the matched item
positions <- c(which(dna_seg$start == 1007350),
                   which(dna_seg$end == 1055793))

positions <- sort(positions) # sorts the positions

  # using the row positions we filter the rows in that range
phage_one <- seg1 <- dna_seg[positions[1]:positions[2], ]
  
  # set color for the dna segs
phage_one$fill <- "#6A3D9A"
phage_one$col <- "#6A3D9A"
  
phage_one_annotation = annotation(text = phage_one$gene,
              x1 = middle(phage_one),
              x2 = middle(phage_one),
              rot = 30)
  
  # plot the DNA segment for phage 1
plot_gene_map(c(phage_one),
                gene_type = "arrows",
                scale = TRUE,
                annotations = phage_one_annotation)

```

#### Phage 2

```{r } 
  
  # region 2

  # which gets the row number of the matched item
  positions <- c(which(dna_seg$start == 1538899),
                   which(dna_seg$end == 1575836))

  positions <- sort(positions) # sorts the positions

  # using the row positions we filter the rows in that range
  phage_two <- seg1 <- dna_seg[positions[1]:positions[2], ]
  
  # set color for the dna segs
  phage_two$fill <- "#6A3D9A"
  phage_two$col <- "#6A3D9A"
  
  phage_two_annotation = annotation(text = phage_two$gene,
              x1 = middle(phage_two),
              x2 = middle(phage_two),
              rot = 30)

  plot_gene_map(c(phage_two),
                gene_type = "arrows",
                scale = TRUE)

```

#### Phage 3

```{r }
  # region 3

  # which gets the row number of the matched item
  positions <- c(which(dna_seg$start == 1781177),
                   which(dna_seg$end == 1787337))

  positions <- sort(positions) # sorts the positions

  # using the row positions we filter the rows in that range
  phage_three <- seg1 <- dna_seg[positions[1]:positions[2], ]
  
  # set color for the dna segs
  phage_three$fill <- "#6A3D9A"
  phage_three$col <- "#6A3D9A"
  
  phage_three_annotation = annotation(text = phage_three$gene,
              x1 = middle(phage_three),
              x2 = middle(phage_three),
              rot = 30)

  plot_gene_map(c(phage_three),
                gene_type = "arrows",
                scale = TRUE)

```

#### Phage 4

```{r }  
    # phage 4

  # which gets the row number of the matched item
  positions <- c(which(dna_seg$start == 1883242),
                   which(dna_seg$end == 1934286))

  positions <- sort(positions) # sorts the positions

  # using the row positions we filter the rows in that range
  phage_four <- seg1 <- dna_seg[positions[1]:positions[2], ]
  
  # set color for the dna segs
  phage_four$fill <- "#6A3D9A"
  phage_four$col <- "#6A3D9A"
  
  phage_four_annotation = annotation(text = phage_four$gene,
              x1 = middle(phage_four),
              x2 = middle(phage_four),
              rot = 30)

  plot_gene_map(c(phage_four),
                gene_type = "arrows",
                scale = TRUE)

```

#### Phage 5

```{r }  
  # which gets the row number of the matched item
  positions <- c(which(dna_seg$start == 2438615),
                   which(dna_seg$end == 2447015))

  positions <- sort(positions) # sorts the positions

  # using the row positions we filter the rows in that range
  phage_five <- seg1 <- dna_seg[positions[1]:positions[2], ]
  
  # set color for the dna segs
  phage_five$fill <- "#6A3D9A"
  phage_five$col <- "#6A3D9A"
  
  phage_five_annotation = annotation(text = phage_five$gene,
              x1 = middle(phage_five),
              x2 = middle(phage_five),
              rot = 30)

  plot_gene_map(c(phage_five),
                gene_type = "arrows",
                scale = TRUE)
```

#### Phage 6

```{r }

   # phage 6

  # which gets the row number of the matched item
  positions <- c(which(dna_seg$start == 3504360),
                   which(dna_seg$end == 3552416))

  positions <- sort(positions) # sorts the positions

  # using the row positions we filter the rows in that range
  phage_six <- seg1 <- dna_seg[positions[1]:positions[2], ]
  
  # set color for the dna segs
  phage_six$fill <- "#6A3D9A"
  phage_six$col <- "#6A3D9A"
  
  phage_six_annotation = annotation(text = phage_six$gene,
              x1 = middle(phage_six),
              x2 = middle(phage_six),
              rot = 30)

  plot_gene_map(c(phage_six),
                gene_type = "arrows",
                scale = TRUE)

```

#### Phage 7

```{r }

  # which gets the row number of the matched item
  positions <- c(which(dna_seg$start == 4458746),
                   which(dna_seg$end == 4507730))

  positions <- sort(positions) # sorts the positions

  # using the row positions we filter the rows in that range
  phage_seven <- seg1 <- dna_seg[positions[1]:positions[2], ]
  
  # set color for the dna segs
  phage_seven$fill <- "#6A3D9A"
  phage_seven$col <- "#6A3D9A"
  
  phage_seven_annotation = annotation(text = phage_seven$gene,
              x1 = middle(phage_seven),
              x2 = middle(phage_seven),
              rot = 30)

  plot_gene_map(c(phage_seven),
                gene_type = "arrows",
                scale = TRUE)

```

#### Phage 8

```{r }  

  # which gets the row number of the matched item
  positions <- c(which(dna_seg$start == 4681892),
                   which(dna_seg$end == 4696380))

  positions <- sort(positions) # sorts the positions

  # using the row positions we filter the rows in that range
  phage_eight <- seg1 <- dna_seg[positions[1]:positions[2], ]
  
  # set color for the dna segs
  phage_eight$fill <- "#6A3D9A"
  phage_eight$col <- "#6A3D9A"
  
  phage_eight_annotation = annotation(text = phage_eight$gene,
              x1 = middle(phage_eight),
              x2 = middle(phage_eight),
              rot = 30)

  plot_gene_map(c(phage_eight),
                gene_type = "arrows",
                scale = TRUE)
```

#### combining the phage regions

```{r }
pdf(file = "./typhi/figures/bacteriophages_1.pdf",
    width = 14)

plot_gene_map(dna_segs =
                list(phage_one, phage_two, phage_three, phage_four,
                     phage_five, phage_six, phage_seven, phage_eight),
                gene_type = "side_blocks",
              dna_seg_labels =
                c("Phage 1", "Phage 2", "Phage 3", "Phage 4", "Phage 5",
                  "Phage 6", "Phage 7", "Phage 8"),
              annotations = list(phage_one_annotation, phage_two_annotation,
                              phage_three_annotation, phage_four_annotation,
                              phage_five_annotation, phage_six_annotation,
                              phage_seven_annotation, phage_eight_annotation),
              annotation_height=2,
              main = "Bacteriophage regions in the isolate AL513382.1",
              dna_seg_scale = TRUE, scale = FALSE)

dev.off()
```


### plot the phaster output for the phage regions
the phage regions are numbered differently for the reference and the isolates.
manually modify the phage regions to match the reference by modifying the summary text file. to determine which regions match which,
there is need to open up the files in the artemis comparison tool and reanotate the text file

#### a function to clean up the phage data
```{r clean_phage_function}
# clean phage data
clean_phage <- 
    function(sample_id) {
    phage_file <-
    file.path(phage_data_location, sample_id, "phage_regions.txt")
    
    phages <-
        read_table(file = phage_file) %>%
    remove_constant() %>%
    remove_empty() %>%
    clean_names() %>%
    filter(!is.na(region_length))

    #split the completeness and score columns
    phages[c("completeness", "score")] <-
        str_split_fixed(phages$completeness_score, "\\(", 2)

    phages$region_length <-
        str_replace_all(phages$region_length, "Kb", "")

    phages$region_length <-
        as.numeric(phages$region_length)

    # get the start position of the bacteriophage
    phages$region_position <-
        str_replace_all(phages$region_position,
        "cluster_002_consensus_polypolish:", "")
    
    phages$region_position <-
        str_replace_all(phages$region_position,
        "cluster_001_consensus_polypolish:", "")
    
    phages[c("start", "stop")] <-
    str_split_fixed(phages$region_position, "-", 2)
    
    phages$start <- 
        as.numeric(phages$start)

    phages$phage_species_num <- 
        as.factor(phages$phage_species_num)

    phages$sample_id <-
        sample_id
    return(phages)
}
```


```{r}
x <- clean_phage(file_names[5])
y <- clean_phage(file_names[1])
z <- clean_phage(file_names[2])
a <- clean_phage(file_names[3])
b <- clean_phage(file_names[4])

all_phages <-
    rbind(x,y,z,a,b)

all_phages$phage_species_num

levels(all_phages$phage_species_num) <-
    as.integer(levels(all_phages$phage_species_num)) %>% sort()

```

```{r}
# save the phages
plotting_dir <- file.path(outdir, Sys.Date())
plot_name <- file.path(plotting_dir, "phage_col_plot.svg")

if (!dir.exists(plotting_dir)) {dir.create(plotting_dir, recursive = TRUE)}

svg(plot_name,
    width = 20,
    height = 6)

plot2 <- all_phages %>%
    mutate(region_position = fct_reorder(region_position, start)) %>%
    ggplot(aes(x = region_position,
        y = region_length, fill = completeness)) +
    geom_col(width = 0.75) +
    theme(axis.text.x = element_text(angle = 30, hjust = 1),
    plot.title = element_text(hjust = 0.5)) +
    ylim(c(0, 60)) +
    scale_fill_viridis_d() +
    xlab("Position on Assembly") +
    ylab("Length Of Bacteriophage (Kb)") +
    labs(fill = "Completeness") +
    facet_wrap(~factor(sample_id,
        c("AL513382.1", "1017142", "A58390", "BKQT8S", "BKQU3X")),
        scales = "free", nrow = 1)

print(plot2)

dev.off() 

```


```{r color_the_phages_by_their_species_number}

# save the phages
plotting_dir <- file.path(outdir, Sys.Date())
plot_name <- file.path(plotting_dir, "phage_species_plot.svg")

if (!dir.exists(plotting_dir)) {dir.create(plotting_dir, recursive = TRUE)}

svg(plot_name,
    width = 20,
    height = 6)

plot3 <- all_phages %>%
    mutate(region_position = fct_reorder(region_position, start)) %>%
    ggplot(aes(x = region_position,
        y = region_length, fill = phage_species_num)) +
    geom_col(width = 0.75) +
    theme(axis.text.x = element_text(angle = 30, hjust = 1),
    plot.title = element_text(hjust = 0.5)) +
    ylim(c(0, 60)) +
    scale_fill_viridis_d() +
    xlab("Position on Assembly") +
    ylab("Length Of Bacteriophage (Kb)") +
    labs(fill = "Phage Species Number") +
    facet_wrap(~factor(sample_id,
        c("AL513382.1", "1017142", "A58390", "BKQT8S", "BKQU3X")),
        scales = "free", nrow = 1)

print(plot3)

dev.off() 
```

```{r}
# save the phages
plotting_dir <- file.path(outdir, Sys.Date())
plot_name <- file.path(plotting_dir, "combined_phage_plot.svg")

if (!dir.exists(plotting_dir)) {dir.create(plotting_dir, recursive = TRUE)}

svg(plot_name,
    width = 20,
    height = 12)

plot2/plot3


dev.off() 
```



# the plasmid vs pHCM1

```{r}

pHCM1 <- "outputs/gbk/pHCM1/pHCM1.gbk"

BKQT8S <- "outputs/gbk/BKQT8S/BKQT8S.gbk"

comparison_file <- "docs/pHCM1-BKQT8S.tsv"

# read the dna segments into genoplotR
segment1 <-
  read_dna_seg_from_file(pHCM1, fileType = "detect")

segment2 <-
  read_dna_seg_from_file(BKQT8S, fileType = "detect")

comparison <- read_comparison_from_blast(comparison_file)

#plot the segments
whole_plasmid <-
  plot_gene_map(dna_segs = list(segment1, segment2),
              comparisons = list(comparison),
              dna_seg_labels = c("pHCM1", "BKQT8S"))
             

#subset the comparison above
dna_seg1 <- filter(segment1, start >= 105911, end <= 187475)
dna_seg2 <- filter(segment2, start >= 64601, end <= 114328)


# write dnaseg 1 for manual annotation
##write_csv(dna_seg1 %>% select(gene, product), "docs/pHCM1_manual_annotation.csv")
#write_csv(dna_seg2 %>% select(gene, product), "docs/BKQT8S_manual_annotation.csv")

dna_seg1_gene_names <- read_csv("docs/pHCM1_manual_annotation.csv")
dna_seg2_gene_names <- read_csv("docs/BKQT8S_manual_annotation.csv")

dna_seg1$gene <- dna_seg1_gene_names$gene

dna_seg2$gene <- dna_seg2_gene_names$gene

# color the is elements
is_elements <- c("IS1X2", "TnAs2" , "tetC" , "IS4321R" , "tetR" , "IS4321L", "ISVsa5" , "ISBcen27", "IS4321R", "IS26", "ISStma11" , "IS5075", "IS1R" , "ISVsa5", "TnAs3", "ISVsa5",  "IS1R" , "IS26", "IS1D")

amr_genes <- c( "dfrA14" , "blaTEM-1", "sul2", "aph(3'')-Ib",  "aph(6)-Id" , "tet(B)", "Sul2", "catA1" )

mercury_resistance <- c("merR",  "merT_1", "merP_1", "merC_1", "merA_1",  "merA_2" , "merC_2" , "merP_2" , "merP_2" , "merT_2" )

other_genes <- c("repE" , "xerC" , "hin" , "hns" , "smc" , "dcm" , "tnpR" , "repA" , "yqjZ" , "umuD" , "umuC", "-", "gltS", "gltS_1", "gltS_2", "hha", "cyaA", "cyaY")

# change colors of blocks for pHCM1
dna_seg1$fill <- 
  case_when(dna_seg1$gene %in% is_elements ~ "#01665e",
            TRUE ~ dna_seg1$fill)

dna_seg1$fill <- 
  case_when(dna_seg1$gene %in% amr_genes ~ "#440154",
            TRUE ~ dna_seg1$fill)

dna_seg1$fill <- 
  case_when(dna_seg1$gene %in% other_genes ~ "#C9C9C9",
            TRUE ~ dna_seg1$fill)


dna_seg1$fill <-
  case_when(dna_seg1$gene %in% mercury_resistance ~ "white",
            TRUE ~ dna_seg1$fill)

# change colors for BKQT8S
dna_seg2$fill <- 
  case_when(dna_seg2$gene %in% is_elements ~ "#01665e",
            TRUE ~ dna_seg2$fill)

dna_seg2$fill <- 
  case_when(dna_seg2$gene %in% amr_genes ~ "#440154",
            TRUE ~ dna_seg2$fill)

dna_seg2$fill <- 
  case_when(dna_seg2$gene %in% other_genes ~ "#C9C9C9",
            TRUE ~ dna_seg2$fill)

dna_seg1$col <- "black"
dna_seg2$col <- "black"

 annotation1 <-
  annotation(text = dna_seg1$gene,
             x1 = middle(dna_seg1),
             rot = 30)
annotation2 <-
  annotation(text = dna_seg2$gene,
             x1 = middle(dna_seg2),
             rot = 30)

#plot the new comparison
plot_gene_map(dna_segs = list(dna_seg1, dna_seg2),
              gene_type = "arrows",
              comparisons = list(comparison),
              #dna_seg_labels = c(dna_seg1 ="pHCM1", dna_seg2 ="BKQT8S"),
              annotations = list( annotation1, annotation2),
              annotation_height = c(2, 1.2),
              annotation_cex = .5,
              dna_seg_scale = TRUE, scale = FALSE,
              main = "Comparison of AMR region in pHCM1 and BKQT8S",
              legend = TRUE,
              dna_seg_label_cex = 1)


```


```{r}
# save the phages
outdir <- "outputs/plots"
plotting_dir <- file.path(outdir, Sys.Date())
plot_name <- file.path(plotting_dir, "plasmid_comparison.pdf")

if (!dir.exists(plotting_dir)) {dir.create(plotting_dir, recursive = TRUE)}

pdf(plot_name,
    width = 40,
    height = 8,
    pointsize = 30)

plot_gene_map(dna_segs = list(dna_seg1, dna_seg2),
              gene_type = "arrows",
              comparisons = list(comparison),
              #dna_seg_labels = c(dna_seg1 ="pHCM1", dna_seg2 ="BKQT8S"),
              annotations = list( annotation1, annotation2),
              annotation_height = c(2, 1.2),
              annotation_cex = .5,
              dna_seg_scale = TRUE, scale = FALSE,
              main = "Comparison of AMR region in pHCM1 and BKQT8S",
              legend = TRUE,
              dna_seg_label_cex = 1)

dev.off() 
```


# load assemblies and comparison files

```{r}
# load assemblies
ct18 <-
  "./results/prokka/chromosome/AL513382.1/AL513382.1.gbk"
s1017142 <-
  "./results/prokka/chromosome/1017142/1017142.gbk"
A53789 <-
  "./results/prokka/chromosome/A53789/A53789.gbk"
A58390 <-
  "./results/prokka/chromosome/A58390/A58390.gbk"
BKQT8S <-
  "./results/prokka/chromosome/BKQT8S/BKQT8S.gbk"
BKQU3X <-
  "./results/prokka/chromosome/BKQU3X/BKQU3X.gbk"
  
# load blast comparisons
comparison_file_1 <-
  "./results/blast/ct18_vs_1017142.txt"
comparison_file_2 <-
  "./results/blast/1017142_vs_A3789.txt"
comparison_file_3 <-
  "./results/blast/A53789_vs_A58390.txt"
comparison_file_4 <-
  "./results/blast/A58390_vs_BKQT8S.txt"
comparison_file_5 <-
  "./results/blast/BKQT8S_vs_BKQU3X.txt"

```
# create and save DNA segments

```{r, eval=FALSE, echo=FALSE}
dna_seg_ref <-
  read_dna_seg_from_genbank(ct18)

dna_seq_1017142 <-
  read_dna_seg_from_genbank(s1017142)

dna_seq_A58390 <-
  read_dna_seg_from_genbank(A58390)

dna_seq_BKQT8S <-
  read_dna_seg_from_genbank(BKQT8S)

dna_seq_BKQU3X <-
  read_dna_seg_from_genbank(BKQU3X)

save(dna_seg_ref,
     dna_seq_1017142,
     dna_seq_A58390,
     dna_seq_BKQT8S,
     dna_seq_BKQU3X,
     file = "data/r/dna_segs.RData")
```



## create and plot Figure4

```{r}
# create dna segments
load("data/r/dna_segs.RData")

dna_seg1 <- filter(dna_seg_ref, start > 3467657, end <= 3475923)
dna_seg1 <- dna_seg1 %>% filter(gene != "hemC") %>% filter( synonym != "STY_03442")
plot_gene_map(list(dna_seg1), gene_type = "arrows", annotations = list(annot1))
annot1 <- annotation(text = dna_seg1$gene, x1=middle(dna_seg1), rot = 30)


# DNA segment two

dna_seg2 <- filter(dna_seq_1017142, start >= 4502005, end <= 4508283)
plot_gene_map(list(dna_seg2), gene_type = "arrows", annotations = list(annot2))
annot2 <- annotation(text = dna_seg2$gene, x1=middle(dna_seg2), rot = 30)

dna_seg2 <- dna_seg2 %>% filter(gene != "hemC")


dna_seg3 <- filter(dna_seq_A58390, start >= 4447227, end <= 4454218)
plot_gene_map(list(dna_seg3), gene_type = "arrows", annotations = list(annot3))

dna_seg3$gene[3] <- "IS1D"
dna_seg3$gene[4] <- "IS1R"
annot3 <- annotation(text = dna_seg3$gene, x1=middle(dna_seg3), rot = 30)

dna_seg3 <- dna_seg3 %>% filter(gene != "hemC") %>% filter(synonym != "STY_04383")



dna_seg4 <- filter(dna_seq_BKQT8S, start >= 4447227, end <= 4454442)

plot_gene_map(list(dna_seg4), gene_type = "arrows", annotations = list(annot4))
dna_seg4$gene[5] <- "IS1D"
dna_seg4$gene[6] <- "IS1R"
annot4 <- annotation(text = dna_seg4$gene, x1=middle(dna_seg4), rot = 30)
dna_seg4 <- dna_seg4 %>% filter(synonym != "STY_04384")



dna_seg5 <- filter(dna_seq_BKQU3X, start >= 4447227, end <= 4454803)
plot_gene_map(list(dna_seg5), gene_type = "arrows", annotations = list(annot5))
dna_seg5$gene[5] <- "IS1D"
dna_seg5$gene[6] <- "IS1R"
annot5 <- annotation(text = dna_seg5$gene, x1=middle(dna_seg5), rot = 30)
dna_seg5 <- dna_seg5 %>% filter(gene != "hemC") %>% filter(synonym != "STY_04381")


# create comparison files
system(command = "blastn -query ./outputs/assemblies/polca/1017142.fasta -subject ./outputs/assemblies/polca/A58390.fasta  -outfmt 6 -out docs/1017142_a58390_comparison.txt")

# comparison file for ct18_a58390
#comparison_1 <-
#  read_comparison_from_blast("docs/ct18_a58390_comparison.txt")

comparison_1 <-
  read_comparison_from_blast(comparison_file_1,
                           sort_by = "per_id")

comparison_2 <-
  read_comparison_from_blast("docs/1017142_a58390_comparison.txt",
                           sort_by = "per_id")
#comparison_3 <-
#  read_comparison_from_blast(comparison_file_3,
#                           sort_by = "per_id")
comparison_4 <-
  read_comparison_from_blast(comparison_file_4,
                           sort_by = "per_id")
comparison_5 <-
  read_comparison_from_blast(comparison_file_5,
                           sort_by = "per_id")



dna_seg1$fill <- 
  case_when(dna_seg1$gene %in% other_genes ~ "#C9C9C9",
            TRUE ~ dna_seg1$fill)

dna_seg3$fill <- 
  case_when(dna_seg3$gene %in% is_elements ~ "#01665e",
            TRUE ~ dna_seg3$fill)

dna_seg3$fill <- 
  case_when(dna_seg3$gene %in% other_genes ~ "#C9C9C9",
            TRUE ~ dna_seg3$fill)


dna_seg4$fill <- 
  case_when(dna_seg4$gene %in% is_elements ~ "#01665e",
            TRUE ~ dna_seg4$fill)

dna_seg4$fill <- 
  case_when(dna_seg4$gene %in% other_genes ~ "#C9C9C9",
            TRUE ~ dna_seg4$fill)


dna_seg5$fill <- 
  case_when(dna_seg5$gene %in% is_elements ~ "#01665e",
            TRUE ~ dna_seg5$fill)

dna_seg5$fill <- 
  case_when(dna_seg5$gene %in% other_genes ~ "#C9C9C9",
            TRUE ~ dna_seg5$fill)


dna_seg1$col <- "black"
dna_seg2$col <- "black"
dna_seg3$col <- "black"
dna_seg4$col <- "black"
dna_seg5$col <- "black"

plot_gene_map(dna_segs = list(dna_seg1, dna_seg2, dna_seg3, dna_seg4, dna_seg5),
              gene_type = "arrows",
              annotations = list(annot1, annot2, annot3, annot4, annot5),
              comparisons = list(comparison_1, comparison_2, comparison_4, comparison_5),
              dna_seg_scale = TRUE,
              dna_seg_labels = c("CT18", "1017142", "A58390", "BKQT8S", "BKQU3X"),
              annotation_height = 1.5,
              annotation_cex = 1,
              main  = "Comparison of the insertion site at cyaA")


```


```{r}
# save the phages
outdir <- "outputs/plots"
plotting_dir <- file.path(outdir, Sys.Date())
plot_name <- file.path(plotting_dir, "Figure4.pdf")

if (!dir.exists(plotting_dir)) {dir.create(plotting_dir, recursive = TRUE)}

pdf(plot_name,
    width = 40,
    height = 18,
    pointsize = 30)

plot_gene_map(dna_segs = list(dna_seg1, dna_seg2, dna_seg3, dna_seg4, dna_seg5),
              gene_type = "arrows",
              annotations = list(annot1, annot2, annot3, annot4, annot5),
              comparisons = list(comparison_1, comparison_2, comparison_4, comparison_5),
              dna_seg_scale = TRUE,
              dna_seg_labels = c("CT18", "1017142", "A58390", "BKQT8S", "BKQU3X"),
              annotation_height = 2,
              main  = "Comparison of the insertion site at cyaA")

dev.off() 
```


## FigureS1

```{r}
# create dna segments

# dna_seg_ref <-
#   read_dna_seg_from_genbank("data/ref/sequence.gb")
# 
# dna_seg_ref <-
#   read_dna_seg_from_genbank(ct18)
dna_seg1 <- filter(dna_seg_ref, start >= 1682538, end <= 1698553)
plot_gene_map(list(dna_seg1), gene_type = "arrows", annotations = list(annot1))
annot1 <- annotation(text = dna_seg1$gene, x1=middle(dna_seg1), rot = 30)



dna_seg2 <- filter(dna_seq_1017142, start >= 2467391, end <= 2483298)
plot_gene_map(list(dna_seg2), gene_type = "arrows", annotations = list(annot2))
annot2 <- annotation(text = dna_seg2$gene, x1=middle(dna_seg2), rot = 30)



dna_seg3 <- filter(dna_seq_A58390, start >= 2551524, end <= 2567431)
plot_gene_map(list(dna_seg3), gene_type = "arrows", annotations = list(annot3))
annot3 <- annotation(text = dna_seg3$gene, x1=middle(dna_seg3), rot = 30)



dna_seg4 <- filter(dna_seq_BKQT8S, start >= 2551470, end <= 2567377)

plot_gene_map(list(dna_seg4), gene_type = "arrows", annotations = list(annot4))
annot4 <- annotation(text = dna_seg4$gene, x1=middle(dna_seg4), rot = 30)



dna_seg5 <- filter(dna_seq_BKQU3X, start >= 2551463, end <= 2567370)
plot_gene_map(list(dna_seg5), gene_type = "arrows", annotations = list(annot5))
annot5 <- annotation(text = dna_seg5$gene, x1=middle(dna_seg5), rot = 30)


# create comparison files
# system(command = "blastn -query ./outputs/assemblies/polca/1017142.fasta -subject ./outputs/assemblies/polca/A58390.fasta  -outfmt 6 -out docs/1017142_a58390_comparison.txt")

# comparison file for ct18_a58390
#comparison_1 <-
#  read_comparison_from_blast("docs/ct18_a58390_comparison.txt")

comparison_1 <-
  read_comparison_from_blast(comparison_file_1,
                           sort_by = "per_id")

comparison_2 <-
  read_comparison_from_blast("docs/1017142_a58390_comparison.txt",
                           sort_by = "per_id")
#comparison_3 <-
#  read_comparison_from_blast(comparison_file_3,
#                           sort_by = "per_id")
comparison_4 <-
  read_comparison_from_blast(comparison_file_4,
                           sort_by = "per_id")
comparison_5 <-
  read_comparison_from_blast(comparison_file_5,
                           sort_by = "per_id")



dna_seg1$fill <- 
  case_when(dna_seg1$gene %in% other_genes ~ "#C9C9C9",
            TRUE ~ dna_seg1$fill)

dna_seg3$fill <- 
  case_when(dna_seg3$gene %in% is_elements ~ "#01665e",
            TRUE ~ dna_seg3$fill)

dna_seg3$fill <- 
  case_when(dna_seg3$gene %in% other_genes ~ "#C9C9C9",
            TRUE ~ dna_seg3$fill)


dna_seg4$fill <- 
  case_when(dna_seg4$gene %in% is_elements ~ "#01665e",
            TRUE ~ dna_seg4$fill)

dna_seg4$fill <- 
  case_when(dna_seg4$gene %in% other_genes ~ "#C9C9C9",
            TRUE ~ dna_seg4$fill)


dna_seg5$fill <- 
  case_when(dna_seg5$gene %in% is_elements ~ "#01665e",
            TRUE ~ dna_seg5$fill)

dna_seg5$fill <- 
  case_when(dna_seg5$gene %in% other_genes ~ "#C9C9C9",
            TRUE ~ dna_seg5$fill)


dna_seg1$col <- "black"
dna_seg2$col <- "black"
dna_seg3$col <- "black"
dna_seg4$col <- "black"
dna_seg5$col <- "black"

plot_gene_map(dna_segs = list(dna_seg1, dna_seg2, dna_seg3, dna_seg4, dna_seg5),
              gene_type = "arrows",
              annotations = list(annot1, annot2, annot3, annot4, annot5),
              comparisons = list(comparison_1, comparison_2, comparison_4, comparison_5),
              dna_seg_scale = TRUE,
              dna_seg_labels = c("CT18", "1017142", "A58390", "BKQT8S", "BKQU3X"),
              annotation_height = 1.5,
              annotation_cex = 1,
              main  = "Comparison of the insertion site at nlpC")


```


```{r}
# save the phages
outdir <- "outputs/plots"
plotting_dir <- file.path(outdir, Sys.Date())
plot_name <- file.path(plotting_dir, "FigureS1.pdf")

if (!dir.exists(plotting_dir)) {dir.create(plotting_dir, recursive = TRUE)}

pdf(plot_name,
    width = 40,
    height = 18,
    pointsize = 30)

plot_gene_map(dna_segs = list(dna_seg1, dna_seg2, dna_seg3, dna_seg4, dna_seg5),
              gene_type = "arrows",
              annotations = list(annot1, annot2, annot3, annot4, annot5),
              comparisons = list(comparison_1, comparison_2, comparison_4, comparison_5),
              dna_seg_scale = TRUE,
              dna_seg_labels = c("CT18", "1017142", "A58390", "BKQT8S", "BKQU3X"),
              annotation_height = 2,
              main  = "Comparison of the insertion site at nlpC")

dev.off() 
```


## Figure s2

```{r}
# create dna segments

dna_seg1 <- filter(dna_seg_ref, start >= 3813583, end <= 3822270)
plot_gene_map(list(dna_seg1), gene_type = "arrows", annotations = list(annot1))
annot1 <- annotation(text = dna_seg1$gene, x1=middle(dna_seg1), rot = 30)


dna_seg2 <- filter(dna_seq_1017142, start >= 8457, end <= 17144)
plot_gene_map(list(dna_seg2), gene_type = "arrows", annotations = list(annot2))
annot2 <- annotation(text = dna_seg2$gene, x1=middle(dna_seg2), rot = 30)


dna_seg3 <- filter(dna_seq_A58390, start >= 8457, end <= 17144)
plot_gene_map(list(dna_seg3), gene_type = "arrows", annotations = list(annot3))
annot3 <- annotation(text = dna_seg3$gene, x1=middle(dna_seg3), rot = 30)



dna_seg4 <- filter(dna_seq_BKQT8S, start >= 8457, end <= 17144)

plot_gene_map(list(dna_seg4), gene_type = "arrows", annotations = list(annot4))
annot4 <- annotation(text = dna_seg4$gene, x1=middle(dna_seg4), rot = 30)


dna_seg5 <- filter(dna_seq_BKQU3X, start >= 8457, end <= 17144)
plot_gene_map(list(dna_seg5), gene_type = "arrows", annotations = list(annot5))
annot5 <- annotation(text = dna_seg5$gene, x1=middle(dna_seg5), rot = 30)


# create comparison files
# system(command = "blastn -query ./outputs/assemblies/polca/1017142.fasta -subject ./outputs/assemblies/polca/A58390.fasta  -outfmt 6 -out docs/1017142_a58390_comparison.txt")

# comparison file for ct18_a58390
#comparison_1 <-
#  read_comparison_from_blast("docs/ct18_a58390_comparison.txt")

comparison_1 <-
  read_comparison_from_blast(comparison_file_1,
                           sort_by = "per_id")

comparison_2 <-
  read_comparison_from_blast("docs/1017142_a58390_comparison.txt",
                           sort_by = "per_id")
#comparison_3 <-
#  read_comparison_from_blast(comparison_file_3,
#                           sort_by = "per_id")
comparison_4 <-
  read_comparison_from_blast(comparison_file_4,
                           sort_by = "per_id")
comparison_5 <-
  read_comparison_from_blast(comparison_file_5,
                           sort_by = "per_id")



dna_seg1$fill <- 
  case_when(dna_seg1$gene %in% other_genes ~ "#C9C9C9",
            TRUE ~ dna_seg1$fill)

dna_seg3$fill <- 
  case_when(dna_seg3$gene %in% is_elements ~ "#01665e",
            TRUE ~ dna_seg3$fill)

dna_seg3$fill <- 
  case_when(dna_seg3$gene %in% other_genes ~ "#C9C9C9",
            TRUE ~ dna_seg3$fill)


dna_seg4$fill <- 
  case_when(dna_seg4$gene %in% is_elements ~ "#01665e",
            TRUE ~ dna_seg4$fill)

dna_seg4$fill <- 
  case_when(dna_seg4$gene %in% other_genes ~ "#C9C9C9",
            TRUE ~ dna_seg4$fill)


dna_seg5$fill <- 
  case_when(dna_seg5$gene %in% is_elements ~ "#01665e",
            TRUE ~ dna_seg5$fill)

dna_seg5$fill <- 
  case_when(dna_seg5$gene %in% other_genes ~ "#C9C9C9",
            TRUE ~ dna_seg5$fill)


dna_seg1$col <- "black"
dna_seg2$col <- "black"
dna_seg3$col <- "black"
dna_seg4$col <- "black"
dna_seg5$col <- "black"

plot_gene_map(dna_segs = list(dna_seg1, dna_seg2, dna_seg3, dna_seg4, dna_seg5),
              gene_type = "arrows",
              annotations = list(annot1, annot2, annot3, annot4, annot5),
              comparisons = list(comparison_1, comparison_2, comparison_4, comparison_5),
              dna_seg_scale = TRUE,
              dna_seg_labels = c("CT18", "1017142", "A58390", "BKQT8S", "BKQU3X"),
              annotation_height = 1.5,
              annotation_cex = 1,
              main  = "Comparison of the insertion site at yidA")


```


```{r}
# save the phages
outdir <- "outputs/plots"
plotting_dir <- file.path(outdir, Sys.Date())
plot_name <- file.path(plotting_dir, "FigureS2.pdf")

if (!dir.exists(plotting_dir)) {dir.create(plotting_dir, recursive = TRUE)}

pdf(plot_name,
    width = 40,
    height = 18,
    pointsize = 30)

plot_gene_map(dna_segs = list(dna_seg1, dna_seg2, dna_seg3, dna_seg4, dna_seg5),
              gene_type = "arrows",
              annotations = list(annot1, annot2, annot3, annot4, annot5),
              comparisons = list(comparison_1, comparison_2, comparison_4, comparison_5),
              dna_seg_scale = TRUE,
              dna_seg_labels = c("CT18", "1017142", "A58390", "BKQT8S", "BKQU3X"),
              annotation_height = 2,
              main  = "Comparison of the insertion site at yidA")

dev.off() 
```

